---
title: "Crossing Simulation Profiling"
author: "Julien Diot"
date: "`r format(Sys.time(), '%Y/%m/%d')`"
colorlinks: true
output:
  html_document:
    self_contained: true
    toc: true
    toc_depth: 4
    toc_float: true
    number_sections: TRUE
    code_folding: hide
    df_print: paged
---

```{r setup, include=FALSE}
knitStartTime <- Sys.time()
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE,
                      include = TRUE,
                      warning = TRUE,
                      message = TRUE,
                      cache = FALSE, # save chunks output
                      results = "hold", # display chunk output in one block
                      
                      # df_print paged options
                      rows.print = 10,
                      max.print = Inf,
                      
                      
                      # format
                      tidy = FALSE, # format code with 'tidy_source()'
                      tidy.opts = list(width.cutoff = 80),
                      strip.white = TRUE, #remove the white lines in the beginning or end of a source chunk
                      
                      fig.align = "center",
                      fig.width = 9,
                      fig.height = 5)

knitr::opts_knit$set(# root.dir = '/tmp', # change wd of all chunks
                     progress = TRUE,
                     verbose = FALSE,
                     width = 80
                     )

```

### Packages / R-options: {.unnumbered}

```{r pkgs_options, results='hide'}
suppressPackageStartupMessages({
  # library(pbapply)
  library(pbmcapply)
  library(parallel)
  library(microbenchmark)
  library(dplyr)
  library(plotly)
})


# load engines' functions
sapply(FUN = source,
       X = list.files("../src",
                      pattern = ".R$",
                      full.names = T))

#  R options
options(max.print = 20)
```


# Introduction

We want to check the time taken by the crossing simulation according to genetic data size and number of crosses.

The variable that can influence the most the simulation time are:

- the genotype size
- the number of different couples
- the number of progeny for each couples
- the chromosomes length in Morgan (the longer, the higher number of crossing over), but this cannot be controlled.




# Profiling

## Load genetic data

```{r loadGenoDta}
phasedGenoFile <- '../data/geno/breedGame_phasedGeno.vcf.gz'
SNPcoordFile <- '../data/SNPcoordinates/breedingGame_SNPcoord.csv'

g <- readPhasedGeno(phasedGenoFile)
SNPcoord <- readSNPcoord(SNPcoordFile)
```

We have at our disposal:

- genotype of `r nrow(g$haplo)` SNPs
- `r ncol(g$haplo)/2` individuals


## Profiling setup

```{r profSetup1}
nCores <- 10 # number of cores on which the simulation are run in parallel
nParams_allCombinations <- 4 # number of different parameters to test for each variables
nParams_aotbe <- 5 # number of different parameters to test for each variables
nRep <- 4 # number of repetition for each scenario

genoSize_max <- nrow(g$haplotypes)
nCouples_max <- 100 # choose(n = ncol(g$haplotypes)/2, 2)
nProgeny_max <- 20

genoSize_min <- genoSize_max / nParams_allCombinations
nCouples_min <- 10 # nCouples_max / nParams_allCombinations 
nProgeny_min <- 10 # nProgeny_max / nParams_allCombinations

profilingSetup <- list(
  genoSize = list(min = genoSize_min,
                  max = genoSize_max),
  nCouples = list(min = nCouples_min,
                  max = nCouples_max), 
  nProgeny = list(min = nProgeny_min,
                  max = nProgeny_max))
```


We will try all combination of `r nParams_allCombinations` different parameters for each variables. (it will generate, `r nParams_allCombinations^3` different scenarios). And also some scenario where 2 variables stay equal (to a medium values) and one have `r nParams_aotbe` different values (it will generate `r nParams_aotbe*3`, additional scenario).

```{r profSetup2}
# create scenarios with all combination of parameters
allCombinationsSetup <- expand.grid(lapply(profilingSetup, function(x){
  round(seq(x$min, x$max, length.out = nParams_allCombinations))
}))


# create scenarios with all other things being equal setup
aotbe_setup <- do.call(rbind, lapply(names(profilingSetup), function(x){
  L <- lapply(names(profilingSetup), function(y){
    if (x == y) {
      round(seq(profilingSetup[[x]]$min,
          profilingSetup[[x]]$max,
          length.out = nParams_aotbe))
    } else {
      round((profilingSetup[[y]]$min + profilingSetup[[y]]$max)/2)
    }
  })
  names(L) <- names(profilingSetup)
  expand.grid(L)
}))


(profilingScenarios <- rbind(allCombinationsSetup, aotbe_setup))

```




## Simulation function

It is hard to use directly the simulation function form the engine because I would need to create input file for each scenario. It will instead use a function similar to the one used in the engine, which will be easier to manage for this profiling. It will use as input the values of the affecting the simulation time and return information about the simulation time. 

> **NOTE: The loading time of phased genotype will not be taken in account in the profiling !**


```{r simFun}
timeSimulation <- function(genoSize, nCouples, nProgeny, haplo, SNPcoord){
  # setup simulation data
  ## Sample SNP's genotype
  selectedSNP <- sample(SNPcoord$SNPid, genoSize)
  haplo <- haplo[selectedSNP, ]
  SNPcoord <- SNPcoord[SNPcoord$SNPid %in% selectedSNP, ]

  ## create crossing table
  allInds <- gsub('_[12]$', '', colnames(haplo)[1:(ncol(haplo)/2)])
  crossTable <- expand.grid(allInds, allInds)
  crossTable <- crossTable[1:(nrow(crossTable)/2),]
  crossTable <- crossTable[sample(nrow(crossTable), nCouples),]
  colnames(crossTable) <- c('ind1', 'ind2')
  crossTable$n <- nProgeny
  crossTable$names <- paste(crossTable$ind1, '_x_', crossTable$ind2)

  # simulation
  t <- Sys.time()
  parentPopulation <- initializeSimulation(haplotypes = haplo,
                                           SNPcoord = SNPcoord)
  simulatedIndividuals <- breedSimulatR::makeCrosses(crosses = crossTable,
                                                     pop = parentPopulation)
  simulatedPop <- breedSimulatR::population$new(inds = simulatedIndividuals,
                                                verbose = FALSE)
  simulatedPop$writeVcf(tempfile(fileext = '.vcf.gz'))
  simTime <- difftime(Sys.time(), t, units = 's')

  return(as.numeric(simTime))
}
```

## Run Profiling

### All combinations
```{r runProf, eval=TRUE}
# repeat the profiling setup
set.seed(2022)
profilingScenarios <- profilingScenarios[rep(seq(nrow(profilingScenarios)), nRep),]
profilingScenarios$rep <- rep(seq(nRep), each = nrow(profilingScenarios)/nRep)
simOrder <- sample(nrow(profilingScenarios)) # randomize simulation order
profilingScenarios <- profilingScenarios[simOrder, ]

# simTimes <- mcmapply(timeSimulation,
simTime <- pbmcmapply(timeSimulation,
                     genoSize = profilingScenarios$genoSize,
                     nCouples = profilingScenarios$nCouples,
                     nProgeny = profilingScenarios$nProgeny,
                     MoreArgs = list(haplo = g$haplotypes,
                                     SNPcoord = SNPcoord),
                     mc.cores = nCores)
if (!is.null(ncol(simTime))) {
  simTime <- t(simTime)
}
results <- cbind(profilingScenarios, simTime)
results <- results[order(simOrder),]
write.csv(results, 'profilingResults.csv', quote = F, row.names = F)
```


# Analyse profiling

```{r readProfDta}
dta <- read.csv('profilingResults.csv')
# dta <- read.csv('profilingResults_64cores_test.csv')
```





```{r plot1}
#plot
plot_ly(type = "scatter",
        mode = "markers",
        data = dta,
        x = ~genoSize,
        y = ~simTime,
        color = ~nCouples,
        symbol = ~nProgeny,
        hoverinfo = 'text',
        text = apply(dta, 1, function(l) {
          paste(names(l), ":", l, collapse = "\n")
        })
)
```


```{r plot2}
plot_ly(mode = "markers",
        data = dta,
        x = ~genoSize,
        y = ~nProgeny,
        z = ~simTime,
        color = ~nCouples)

```


```{r linReg1}
# Simple linear reggression
mod <- lm(simTime ~ genoSize + nCouples + nProgeny, data = dta)
summary(mod)
## residual check
par(mfrow = c(2,2))
plot(mod)
```

```{r linReg2}
# Simple linear reggression on squared variables
for (varOfInterest in c('genoSize', 'nCouples', 'nProgeny')) {
  dta[, paste0(varOfInterest, 2)] <- dta[, varOfInterest]^3
}

mod <- lm(simTime ~ genoSize2 + nCouples2 + nProgeny2, data = dta)
summary(mod)
## residual check
par(mfrow = c(2,2))
plot(mod)



predDta <- data.frame(genoSize = 10000, nCouples = 500, nProgeny = 300)
for (varOfInterest in c('genoSize', 'nCouples', 'nProgeny')) {
  predDta[, paste0(varOfInterest, 2)] <- predDta[, varOfInterest]^3
}
predictedTime <- predict(mod, predDta)
predictedTime <- as.difftime(tim = predictedTime, units = 'secs')
as.numeric(predictedTime, units = "hours")
```

# Appendix {.unnumbered}

<details>
<summary style="margin-bottom: 10px;">Session Information (click to expand)</summary>
<!-- Place an empty line before the chunk ! -->

```{r sessionInfo, echo=FALSE}
  options(max.print = 10000)
  cat("Document generated in:\n")
  print(Sys.time() - knitStartTime)
  if (Sys.info()["sysname"] == "Linux") {
    cat("\nCPU: ")
    cat(unique(system("awk -F': ' '/model name/{print $2}' /proc/cpuinfo", intern = T)))
    cat("\nMemory total size: ")
    cat(as.numeric(system("awk '/MemTotal/ {print $2}' /proc/meminfo", intern = T))*10^(-6), "GB")
  }
  cat("\n\n\nSession information:\n")
  print(sessionInfo(), locale = FALSE)
```

</details>
