---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# Data

This folder contain example data to run this engine.

## `geno/`

Genetic data.

The `.vcf` files are not gitted.

```{r}
list.files("geno")
```

```{r}
cat(paste(readLines("geno/testMarkerData01.vcf", n=20), collapse = "\n"))
```


## `pheno/`

Phenotypic data.

```{r}
list.files("pheno")
```

```{r}
cat(paste(readLines("pheno/testPhenoData01.csv", n=10), collapse = "\n"))
```


## `pedigree/`

Example of pedigree file. This is fictive pedigree file inspired by Greek 
mythology. It is the same pedigree in 3 versions:

- Individual name with character string
- Individual name with numbers
- Pedigree with missing founders

```{r}
cat(paste(readLines("pedigree/testPedData_char.csv", n=10), collapse = "\n"))
```

## `results/`

To store the example of results created by the engine.

```{r}
list.files("results")
```

```{r}
cat(paste(readLines("results/gwasResult.json", n=30), collapse = "\n"))
```

```{r}
cat(paste(readLines("results/pedigreeRelationship.csv", n=30), collapse = "\n"))
```
```{r}
cat(paste(readLines("results/pedigreeRelationship.json", n=30), collapse = "\n"))
```

# Reference

The genotypic and phenotypic data used as example come from:

Keyan Zhao, Chih-Wei Tung, Georgia C. Eizenga, Mark H. Wright, M. Liakat Ali, Adam H. Price, Gareth J. Norton, M. Rafiqul Islam, Andy Reynolds, Jason Mezey, Anna M. McClung, Carlos D. Bustamante & Susan R. McCouch (2011). [Genome-wide association mapping reveals a rich genetic architecture of complex traits in *Oryza sativa*.](http://www.nature.com/ncomms/journal/v2/n9/full/ncomms1467.html) Nat Comm 2:467 | DOI: 10.1038/ncomms1467, Published Online 13 Sep 2011.
